---
title: "Chapter 2 HW"
author: "Ulices Gonzalez"
date: "Febuary 13th, 2025"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

* * *

#### Exercise 2: Explain whether each scenario is a classifcation or regression problem, and indicate whether we are most interested in inference or prediction. Finally, provide n and p.

(a) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.

Regression - This would be a regression as we are looking at numerical values such as profit and number of employees, infer the factors that affect the salary of the CEO.  
Inference - Using these values indicates that there is interest in coming up with an inference using the data.  
n = 500 ; p = 3;

(b) We are considering launching a new product and wish to know whether it will be a success or a failure. We collect data on 20 similar products that were previously launched. For each product we have recorded whether it was a success or failure, price charged for the product, marketing budget, competition price, and ten other variables.

Classification - Since it is mentioned that the purpose of the collection of data being used to determine if a product will be a success or a failure it would mean that this would be a classification.  
Prediction - This would then mean that the data collected would be used to provide a prediction for this classification.  
n = 20; p = 13;

(c) We are interested in predicting the % change in the USD/Euro exchange rate in relation to the weekly changes in the world stock markets. Hence we collect weekly data for all of 2012. For each week we record the % change in the USD/Euro, the % change in the US market, the % change in the British market, and the % change in the German market.

Regression - For this last problem due to the use of numerical values to make a predication in the % change of the USD/Euro Exchange rate this problem would be a regression problem.  
Prediction - It is also explained to us that the intention of the problem is to come up with a prediction.  
n = 52; p = 3;


#### Exercise 4: You will now think of some real-life applications for statistical learning.

(a) Describe three real-life applications in which classification might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your
answer.

Garbage Day Pickup  
Response - Will the garbage be picked up?  
Predictors - Day of the Week, Time of Day, Is it a Holiday?  
Goal - To infer if your Garbage will picked up.  
If it is a holiday typically garbage is not picked up so if it is not a holiday you can consider whether it is the assigned day of the week for pick up along with the time of the day as it will usually only be done form the morning to early afternoon but not after.

Tax Filing Eligibility  
Response - Can I file my taxes?  
Predictors - W2 has been issued, Tax Appointment Scheduled, Online Tax Process Started, Has 4/15 passed?  
Goal - To infer whether you are able to file your taxes either online or in person.  
Almost every year your taxes need to be filed by the 15th of April or whichever day is set for the year. If the day has not passed then you can file your taxes only if you have your W2 and from there you either need to have an appointment set up or the process started online. Otherwise you wouldn't be able to file your taxes.

Incoming Paycheck/Direct Deposit  
Response - Am I going to get paid?  
Predictors - Timecard Submission, Check/Deposit Received, Has it been two weeks since last check/deposit?  
Goal - To predict whether your Check/Deposit has been recieved or is in process.  
For most employees their payment is set up biweekly either for them to recieve a check or a deposit in their bank account. If your timecard has been submitted for the last pay period and its been two weeks or more either your payment is in process or it has already been receieved.

(b) Describe three real-life applications in which regression might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer.

Temperature Forecast for a Given Day  
Response - What will be the highest and lowest temperature during the day?  
Predictors - Last Week's Highest/Lowest Daily Temperatures, High/Low from same day year before  
Weather and temperature forecasts can change regularly but by using the highs and lows from the past 7 days you can predict probable highs and lows of the current day or the next using the average taken. This can then be verified by looking at the temperature on the same day the year before to predict a probable range.

Car Resale Value  
Response - How much can a car sell for?  
Predictors - Car Age, Mileage, Previous Owner(s), Fuel Efficiency, Make/Model, Condition, Prior Accidents  
Goal - To infer the future resell value of a car.  
Overtime a car can depreciate in value depending based on different factors, if a car has low fuel efficiency yet has high mileage if it is a recent car then its likely to have a lower resale value than a car with better fuel efficiency and mileage. The car manufacturer and type come also can affect this value either positively or negative. Lastly if there is a history of accidents or many prior owners this could negatively impact the resale value. When considered altogether it is then possible to infer the resale value of a car.

Life Insurance Policy Value  
Response - What will the value of a life insurance policy be?  
Predictors - Age, Medical History, Lifestyle  
Goal - To infer total value of a customers life insurance policy.  
When someone applies for a life insurance policy insurance agents often will ask questions regarding the customers health. One's medical history can lead to a lower or higher value depending on what is found. Lifestyle choices are also taken into account as someones weight, history of drinking and or smoking can lead to health issues down the line that may decrease the value of the policy. Lastly the younger a customer is the higher their potential policy value will be so long as their medical history and lifestyle are not high risk. Using these it is possible for both customers and agents to infer the value of a given policy before anything is set.

(c) Describe three real-life applications in which cluster analysis might be useful.

Diet Preferences  
If a restaraunt is trying to develop a new menu they can group customers based off what their meals of choice are when they come in. It allows for customers to be grouped based on if they eat meat, if they're pescatarian, vegetarian, vegan, and other dietary preferences. This then allows a restaraunt owner to revamp their menu or make a new one altogether with new options that fit each group or cluster.

Book Club Selections  
Typically a book club will sit down and select a single book to be the focus of their meeting but because of how diverse the tastes are from each participant creating clusters based on preferred genre can allow for smaller discussions to be part of a book club if they could not come to a final say on a single book. This would allow the hosts of the book club to set out different groups such as science fiction, mystery, western, biography, and many other options that can be used. This allows everyone to participate while also making it easier in cases where some didnt have access or just didnt read a single book.

Museum Tour Groups  
Depending on the museum you may got to there will be a lot of variety when it comes to the exhibits present. In cases where guests have a preference on the tour and exhibits they take you can cluster them based off a giving focus. An example of this would be the legion of honor in the city, they have classical paintings, contemporary, sculptures, photpgrahy and much more. Clusters like these would make it easier for field trips, family excursions, and other activities to suit the preferences of people in the museum while also filling the capacity for a given tour.

* * *

## Applied Questions

#### Exercise 8: This exercise relates to the College data set. It contains a number of variables for 777 different universities and colleges in the US.

(a)

```{r}
college <- read.csv("DataSets/College.csv")
```

(b)

```{r}
rownames(college) <- college[, 1]
View(college)

college <- college[, -1]
View(college)
```

(c)
i.  
```{r}
summary(college)

```

ii.  
```{r}
college$Private <- as.factor(college$Private)
pairs(college[, 1:10])
```

iii.  
```{r}
plot(college$Private, college$Outstate, xlab = "Private", ylab = "Outstate", main = "Private vs Outstate")
```

iv.  
```{r}
Elite <- rep("No", nrow(college))
Elite[college$Top10perc > 50] <- "Yes"
Elite <- as.factor(Elite)
college <- data.frame(college, Elite)

summary(Elite)

plot(Elite, college$Outstate, xlab = "Elite", ylab = "Outstate", main = "Elite vs Outstate")

```

v.  
```{r}
par(mfrow = c(2, 2))
for (n in c(5, 25, 50, 100)) 
{
  hist(college$Accept, breaks = n, xlab = "Students", main = paste("Acceptances: n = ", n))
}
```

vi.  
Based on the given data majority of schools included are Private. Looking at the acceptances there seems to be more students enrolled at a given time if total enrollments for a university are between 0 to 5000 students but skews towards higher acceptance rates towards 0 to 1250 approx. As the number of student grows there is a decline in acceptances amongst all universities in the dataset.


#### Exercise 10: This exercise involves the Boston housing data set.

(a) 
```{r}
library(ISLR2)
```

The data set contains 506 rows which represent Suburbs of Boston and has 13 columns which represent variables such as number of rooms, median value, and per capita crime rate for example.  

(b) 
```{r}
pairs(Boston)
```
The relationship with the Charles River Dummy Variable is shown to have a constant relationship throughout the data set which makes sense since it can either be 0 or 1 but instead of varying significantly in its relationship with the rest of the variables almost all the graphs show the same relationship.

(c)
Looking at the scatterplots there seems to be an increase in crime per capita the older the neighborhood is. A similar relationship is seen with higher access to highways, higher tax rate, and higher pt ratio correlating with a higher crime per capita rate.

(d)
```{r}
par(mfrow = c(1, 3))
hist(Boston$crim, breaks = 25, xlab = "Suburbs", ylab = "Crime per Capita")
hist(Boston$tax, breaks = 25, xlab = "Suburbs", ylab = "Tax Rate per $10,000")
hist(Boston$ptratio, breaks = 25, xlab = "Suburbs", ylab = "Pupil Teach Ratio")
```

When looking at the range present in Crime rates theres a large gap as there are only a few suburbs with extremely high crime rates while most others have significantly lower. The range while large is smaller with Tax Rates in suburbs with most suburbs being between below 80. Lastly when looking at the pupil to teacher ratio theres a much smaller range.

(e)
```{r}
sum(Boston$chas)
```

There are 35 suburbs near the Charles River

(f)
```{r}
summary(Boston$ptratio)
```

The median for Pupil-Teach Ration is 19.05

(g)
```{r}
lowMed <- Boston[order(Boston$medv),]
lowMed[1,]
summary(Boston)
```
When comparing the vaues for the 399th suburb which has the lowest median value while it has a lower crime rate has a higher than average nitrogen oxide concentration, less rooms present, and is one of the oldest suburbs in Boston. Due to it skewing past the average and into the third quartile for most of the variables it seems to be the most affected by all of them.

(h)
```{r}
subset(Boston, Boston$rm > 7)
summary(subset(Boston, Boston$rm > 7))
subset(Boston, Boston$rm > 8)
summary(subset(Boston, Boston$rm > 8))
```
There is a total of 64 suburbs with homes that have more than 7 rooms, 13 with more than 8. These suburbs tend to have a significantly lower crime rate then almost all others in the data set along with some of the higher propertly values.

